Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-06-16 trunk/paper_enahpc14/Conclusion.tex
updates paper according to page proofs

We  have   presented  a  methodology  for   comparing  performance  of
\cosmoart,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandy Bridge (2.6 GHz) system has
lower time-to-solution but higher energy-to-solution than an Intel Ivy
Bridge (2.2 GHz) system, although  the two metrics are indeed strongly
correlated.   On the  other  hand, the  usage  of energy-friendly  MPI
waiting techniques analyzed  with our power-per\-for\-man\-ce framework over
a series of  runs of \cosmoart on \tinto  and power traces demonstrate
that   it  is   possible  to   reduce  both   power   dissipation  and
energy-to-solution   while  maintaining   (or  even   decreasing)  the
time-to-solution. The resulting profiles indicate that simple changes,
such as making use of the  \emph{degraded} MPI policy rather than the \emph{aggressive}
is  possible to  slightly  reduce both  power  dissipation and  energy
consumption (by 2\,\%).

>>>>

We  have   presented  a  methodology  for   comparing  performance  of
\cosmoart,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: a system based on Intel Xeon Sandy Bridge
(2.6  GHz) has  lower time-to-solution  but  higher energy-to-solution
than on a platform based on  Intel Xeon Ivy Bridge (2.2 GHz), although
the two  metrics are indeed  strongly correlated.  On the  other hand,
the usage of energy-friendly  MPI waiting techniques analyzed with our
power-per\-for\-man\-ce framework  demonstrate that it  is possible to
reduce the  energy-to-solution while maintaining  (or even decreasing)
the  time-to-so\-lu\-tion. The resulting  power profiles  of \cosmoart
indicate   that    simple   changes,    such   as   making    use   of
the \emph{degraded}  MPI policy rather than  the \emph{aggressive} (by
default)  is  possible  to  slightly  reduce  both  power  and  energy
consumption (by 2\,\%).

### 2014-06-16 trunk/paper_enahpc14/Sec3:Environment-setup.tex
updates paper according to page proofs

\begin{itemize}
\item \monch is  a 10-rack NEC-provided clus\-ter at  CSCS composed by
  384   nodes  grouped  in  312  standard  compute  nodes,  24
  large-memory compute nodes,  24 huge-memory nodes, 16 administration
  nodes and 8 servers to provide I/O to Lustre.
  % and  utilized  by  members  of  the  Swiss  Platform  for  Advanced
  % Scientific Computing (PASC) project~\cite{PASC}.  
Each standard  compute node  is comprised of  two Intel Xeon Ivy Bridge EP
E5-2660v2  ten-core processors  operating at  2.2\,GHz,  equipped with
32\,GB  of DDR3 1600\,MHz  RAM and  connected via  InfiniBand Mellanox
SX6036 and FDR switches.  For our experiments, only a subset of 52 was used. %
% only a full rack constituted exclusively of 52 nodes was used.
  % \monch is  slated to stay  in service without hardware  upgrade for
  % the  duration  of  the  Exa2Green project,  allowing  an  identical
  % configuration  to  be  recreated  for  future  assessments  of  the
  % baseline.

\item \pilat is a cluster at CSCS composed of 42 compute nodes.  
  % and used as Piz Daint pre-post processing cluster. The 2 login and
  % 42  compute  nodes  are   consolidated  into  11  twin-pair  Intel
  % E5-Series  DALCO  r2264i4t   2U  scalable  compute  modules,  each
  % containing 4 compute nodes based on two ...
  Each of them is comprised of two Intel Xeon Sandy Bridge EP E5-2670 eight-core processors operating
  at 2.6\,GHz equipped with 64\,GB of DDR3 1600\,MHz RAM and connected
  via  InfiniBand   Mellanox  SX6036   and  FDR  switches.    For  our
  experiments, a subset constituted of 42 compute  nodes was used.
  % \pilat   is   based   on   Intel's   second-generation   processors,
  % conventional in  HPC systems  and known to  be more  power consuming
  % than its Ivy Bridge successor.

@@

The \cosmoart  baseline is a pure MPI-based  Fortran~90 code currently
running on distributed multi-core systems only.  The software stack on
the platforms  was controlled using the modules  framework which gives
an  easy and  flexible  mechanism to  access  to all  of the  provided
compilers, tools  and applications.  For our  initial benchmarking, we
opted for the GNU compiler GCC 4.8.1 on \monch and GCC 4.8.2 on \pilat
and  \tinto)  using  the \texttt{-O3}  compiler  flag  in  favor of  the  Intel
compilers 14.0.1, which  delivered inferior performance.  In addition,
we  installed the following  implementations of  MPI: MVAPICH2  1.9 on
\monch and \pilat, and OpenMPI 1.6.5  on \tinto. On the other hand, we
used HDF5  1.8.12 and NetCDF  (4.3.1) libraries for the  management of
extremely large and complex data collections.  All computes nodes have
an     operating    system     based     on    GNU/Linux     featuring
``2.6.32-358.11.1.el6'' kernel  in \monch, ``3.0.101-0.15-default'' in
\pilat and ``2.6.18-238.19.1.el5'' in \tinto.

>>>>

\begin{itemize}
\item \monch is  a 10-rack NEC-provided clus\-ter at  CSCS composed by
  384   nodes  grouped  in  312  standard  compute  nodes,  24
  large-memory compute nodes,  24 huge-memory nodes, 16 administration
  nodes and 8 servers to provide I/O to Lustre.
  % and  utilized  by  members  of  the  Swiss  Platform  for  Advanced
  % Scientific Computing (PASC) project~\cite{PASC}.  
Each standard  compute node  is comprised of  two Intel Xeon Ivy Bridge EP
E5-2660v2  ten-core processors  operating at  2.2\,GHz,  equipped with
32\,GB  of DDR3 1600\,MHz  RAM and  connected via  InfiniBand Mellanox
SX6036 and FDR switches.  For our experiments, a full rack  of 52 standard compute nodes was used. %
% only a full rack constituted exclusively of 52 nodes was used.
  % \monch is  slated to stay  in service without hardware  upgrade for
  % the  duration  of  the  Exa2Green project,  allowing  an  identical
  % configuration  to  be  recreated  for  future  assessments  of  the
  % baseline.

\item \pilat is a cluster at CSCS composed of 42 compute nodes.  
  % and used as Piz Daint pre-post processing cluster. The 2 login and
  % 42  compute  nodes  are   consolidated  into  11  twin-pair  Intel
  % E5-Series  DALCO  r2264i4t   2U  scalable  compute  modules,  each
  % containing 4 compute nodes based on two ...
  Each of them is comprised of two Intel Xeon Sandy Bridge EP E5-2670 eight-core processors operating
  at 2.6\,GHz equipped with 64\,GB of DDR3 1600\,MHz RAM and connected
  via  InfiniBand   Mellanox  SX6036   and  FDR  switches.    For  our
  experiments, a full rack constituted of 42 compute  nodes was used.
  % \pilat   is   based   on   Intel's   second-generation   processors,
  % conventional in  HPC systems  and known to  be more  power consuming
  % than its Ivy Bridge successor.

@@

The \cosmoart  baseline is a pure MPI-based  Fortran~90 code currently
running on distributed multi-core systems only.  The software stack on
the platforms  was controlled using the modules  framework which gives
an  easy and  flexible  mechanism to  access  to all  of the  provided
compilers, tools  and applications.  For our  initial benchmarking, we
opted for the GNU compiler GCC 4.8.1 on \monch and GCC 4.8.2 on \pilat
and \tinto) using the \texttt{-O3} compiler flag in favor of the Intel
compilers 14.0.1, which  delivered inferior performance.  In addition,
we  installed the following  implementations of  MPI: MVAPICH2  1.9 on
\monch and \pilat, and OpenMPI 1.6.5  on \tinto. On the other hand, we
used  HDF5 1.8.12  and NetCDF  4.3.1 libraries  for the  management of
extremely large and complex data collections.  All computes nodes have
an     operating    system     based     on    GNU/Linux     featuring
``2.6.32-358.11.1.el6'' kernel  in \monch, ``3.0.101-0.15-default'' in
\pilat and ``2.6.18-238.19.1.el5'' in \tinto.
